\documentclass{article}
\usepackage{amsmath,amsthm}
\usepackage{mathtools}

\title{AM-GM Inequality Done Right}
\author{Sebastian Miles}
\date{}

\begin{document}
	\maketitle
	\section{Statement}
	Given the positive reals $a_1,a_2, \cdots, a_n$ the following inequality holds
	\begin{align*}
		\frac{a_1+a_2+ \cdots + a_n}{n} \geq \sqrt[n]{a_1a_2 \cdots a_n}
	\end{align*}
	with the equality if and only if $a_1=a_2=\cdots=a_n$.
	
	\section{Cauchy's Forward-backward induction proof}
	The FB (Forward-backward) induction is a rather rare variant of proof by induction based on the idea of first proving the base case as usual such as $P(2)$. Then proving for arbitrary large n such that if $P(n)$ is true then $P(2n)$ which follows recursively that $P(2^k)$ must be true. The final step is to show that $P(n) \Rightarrow P(n-1)$ which will prove that $P(n)$ holds for any $n \geq 2$.
	
	\begin{quote}
		\textit{i}). Base case.\\
		\textit{ii}). $P(n) \Rightarrow P(2n)$.\\
		\textit{iii}). $P(n) \Rightarrow P(n-1)$.
	\end{quote}
	\noindent
	Step \textit{i}): The base case is when $n=2$ because this is the smallest n for which there is an inequality, $n=1$ would give us an equality. Given $n=2$, the AM-GM inequality implies:
	\begin{equation}
		\frac{a+b}{2} \geq \sqrt{ab}
		\label{basecasen2}
	\end{equation}
 	\noindent
	To prove this we square both sides which yields
	\begin{align*}
		\frac{a^2+2ab+b^2}{4} \geq ab &\iff a^2+2ab+b^2 \geq 4ab& \\
		a^2-2ab+b^2 \geq 0 &\iff (a-b)^2 \geq 0.&
	\end{align*}
	Because $(a-b)^2 \geq 0$ is obviously true, we conclude that
	$\frac{a+b}{2} \geq \sqrt{ab}$
	must be true. \qed
	\newline
	Step \textit{ii}): To perform this step, we first suppose $P(n)$ is true then we prove that $P(2n)$ must be true. Suppose for positive reals $a_1,a_2, \cdots, a_n$ that
	\begin{align*}
		\frac{a_1+a_2+ \cdots + a_n}{n} \geq \sqrt[n]{a_1a_2 \cdots a_n}
	\end{align*}
	holds. Then equivalently we have
	\begin{equation}
		a_1+a_2+ \cdots + a_n \geq n\sqrt[n]{a_1a_2 \cdots a_n}
		\label{cauchyp2n}
	\end{equation}
	We start with the sum $a_1+\cdots+a_{2n}$ and then prove for $2n$:
	\begin{align*}
		a_1+\cdots+a_n+a_{n+1}+\cdots+a_{2n} \geq n\sqrt[n]{a_1\cdots a_n} + n\sqrt[n]{a_{n+1}\cdots a_{2n}}
	\end{align*}
	Here we use the fact that the sum can be split up in to two parts $a_1 + \cdots + a_n$ and $a_{n+1} + \cdots + a_{2n}$, then we apply the inequality from equation \ref{cauchyp2n} on each part. Recall the base case inequality (\ref{basecasen2}), this can be slightly modified to the form $a+b \geq 2\sqrt{ab}$ which we can apply directly to the inequality above.
	\begin{align*}
		n\sqrt[n]{a_1\cdots a_n} + n\sqrt[n]{a_{n+1}\cdots a_{2n}} \geq 2\sqrt{n^2\sqrt[n]{a_1\cdots a_{2n}}} = 2n\sqrt[2n]{a_1\cdots a_{2n}}
	\end{align*}
	This means that we have the following inequality
	\begin{align*}
		a_1 + \cdots + a_{2n} \geq 2n\sqrt[2n]{a_1\cdots a_{2n}}
	\end{align*}
	We have now shown that if 
	$$\frac{a_1+ \cdots + a_n}{n} \geq \sqrt[n]{a_1\cdots a_n}$$ 
	then 
	$$\frac{a_1+ \cdots + a_{2n}}{2n} \geq \sqrt[2n]{a_1\cdots a_{2n}}.$$ \qed
	\newline
	Step \textit{iii}): Suppose for positive reals $a_1,a_2, \cdots, a_n$ that
	\begin{equation}
		\frac{a_1+a_2+ \cdots + a_n}{n} \geq \sqrt[n]{a_1a_2 \cdots a_n}
		\label{step3assumption}
	\end{equation}
	To prove P(n-1) we start with $\frac{a_1+\cdots +a_{n-1}}{n-1}$ and then prove the right side of the inequality. In order to perform the induction we must somehow turn the expression in to a form such that we can apply inequality \ref{step3assumption}. To do this we may let $$\frac{a_1+\cdots +a_{n-1}}{n-1}=\frac{a_1+\cdots +a_n}{n}$$ 
	and then solve for $a_n$. This means that we can express $\frac{a_1+\cdots +a_{n-1}}{n-1}$ in a way such that we can use the induction assumption (\ref{step3assumption}).
	\begin{align*}
		&\frac{a_1+\cdots +a_{n-1}}{n-1}=\frac{a_1+\cdots +a_n}{n} \\
		\iff& n(a_1+\cdots +a_{n-1})=(n-1)(a_1+\cdots +a_n) \\
		\iff& \frac{a_1+\cdots +a_{n-1}}{n-1}=a_n
	\end{align*}
	Because we have $$\frac{a_1+\cdots +a_{n-1}}{n-1}=\frac{a_1+\cdots +a_n}{n}$$ when $\frac{a_1+\cdots +a_{n-1}}{n-1}=a_n$ it means that we have
	\begin{align*}
		\frac{a_1+\cdots +a_{n}}{n} &\geq \sqrt[n]{a_1\cdots a_n} \\
		\frac{a_1+\cdots +a_{n-1}}{n-1} &\geq \sqrt[n]{a_1\cdots a_{n-1} \cdot \frac{a_1+\cdots +a_{n-1}}{n-1}}\\
		\left(\frac{a_1+\cdots +a_{n-1}}{n-1}\right)^n &\geq a_1\cdots a_{n-1} \cdot \frac{a_1+\cdots +a_{n-1}}{n-1} \\
		\left(\frac{a_1+\cdots +a_{n-1}}{n-1}\right)^{n-1} &\geq a_1\cdots a_{n-1}\\
		\frac{a_1+\cdots +a_{n-1}}{n-1} &\geq \sqrt[n-1]{a_1\cdots a_{n-1}}
	\end{align*}
	Which proves that $n-1$ must also be true. \qed
	\newline
	We have now completed the forward-backward induction proof which proves the AM-GM inequality for $n \geq 2$, the $n=1$ case is trivial as $\frac{a}{1}=a^1$, thus we have proven that for positive reals $a_1,a_2, \cdots, a_n$:
	
	\[ \frac{a_1+a_2+ \cdots + a_n}{n} \geq \sqrt[n]{a_1a_2 \cdots a_n}  \tag*{for $n \geq 1$}   \]
	\qed
\end{document}